---
title: "HoldOutMetabric"
output: html_document
date: "2025-04-20"
---

## Concordance index multiverse.

Load libraries

```{r, include=FALSE}
# Survival metrics
library(reticulate)
library(arrow)
library(caret)
library(riskRegression)
library(prodlim)
library(pec)
library(survival)
library(rhdf5)
library(randomForestSRC)
library(survAUC)
library(Hmisc)
library(dplyr)
# Plotting
library(gridExtra)
# Parallelization
library(doFuture)
library(future)
library(progressr)
library(foreach)
library(MASS)
library(flexsurv)
library(furrr)
library(pysurvivalR)
library(survivalmodels)
library(mlr3)
```


```{r setup, include=FALSE}
#Sys.unsetenv("RETICULATE_PYTHON")
Sys.setenv(OMP_NUM_THREADS = "1")       # Limits OpenMP to 1 thread
Sys.setenv(NUMBA_NUM_THREADS = "1")     # Limits Numba to 1 thread
Sys.setenv(MKL_NUM_THREADS = "1")       # Limits Intel MKL to 1 thread
Sys.setenv(KMP_WARNINGS = "0")          # Disables OpenMP warnings
Sys.setenv(OPENBLAS_NUM_THREADS = "1")  # Limits OpenBLAS to 1 thread

# #Sys.setenv(NUMBA_DISABLE_JIT = "0")  
# library(reticulate)

#use_condaenv("/opt/homebrew/Caskroom/miniforge/base/envs/py-rstudio", required=TRUE)
#use_virtualenv("~/.virtualenvs/venv-DeSurv_python_R")
#use_python("/opt/homebrew/Caskroom/miniforge/base/envs/venv-DeSurv3/bin/python3.9")
#py_config()
```


```{r}
# Dataset from Deep Surv, also used in DeSurv:
h5_test <- h5read("./Datasets/metabric_IHC4_clinical_train_test.h5", 
                  name="test")

h5_train <- h5read("./Datasets/metabric_IHC4_clinical_train_test.h5",
                   name="train")

# Make a dataset for training:
train <- data.frame(t(h5_train$x))
train$status <- h5_train$e
train$time <- h5_train$t

# Make a dataset for test:
test<- data.frame(t(h5_test$x))
test$status <- h5_test$e
test$time <- h5_test$t

# Order
test <- test[order(test$time, -test$status),]
train <- train[order(train$time, -train$status),]
test$time <- test$time + 1e-8
train$time <- train$time + 1e-8

mb <- rbind(train, test) # for crossvalidation settings

```

```{r}
# read the processed file
mb <- read.table("./Datasets/metabric_preprocess_multiverse.csv", sep="," , header = TRUE)

# Keep the id as index
rownames(mb) <- mb$PATIENT_ID

# Remove the id column
mb <- mb[,2:12]

# Sote the previous colnames to keep track of them
cols <- colnames(mb)

# Asign new column names
newcols <- c("X1", "X2", "X3", "X4","X5", "X6", "X7", "X8", "X9", "time", "status")
colnames(mb) <- newcols

# NO NEED OF SUMING, there is no 0s... 
min(mb$time)
#mb$time <- mb$time + 1e-8

# Order according to time and status
mb[order(mb$time, -mb$status),]

```

```{r}
source("./CindexHelperFunctions.R")
```


### HOLDOUT and BOOTSTRAP


#### DeepSurv test/train splits:


For the hyperparameters of deep learning models that have been trained with survivalmodels package is relevant to look at the pytorch optimizer: https://raphaels1.github.io/survivalmodels/reference/get_pycox_optim.html 

Also the parameters for each model:
1) Deephit: https://raphaels1.github.io/survivalmodels/reference/deephit.html
2) DeepSurv: https://raphaels1.github.io/survivalmodels/reference/deepsurv.html
3) CoxTime: https://raphaels1.github.io/survivalmodels/reference/coxtime.html

\begin{table}[ht]
\centering
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Hyper-parameter}     & \textbf{DeepSurv} & \textbf{CoxTime} & \textbf{DeepHit}  & RandomForest \\ \midrule
Optimizer                    & adam              & adam             & adam              & NA \\
Activation                   & selu              & relu             & relu              & NA\\
\# Dense Layers              & 1                 & 2                & 2                 & - \\
\# Nodes / Layer             & 41                & 32, 32           & 32, 32            & 100 trees\\
Learning Rate                & 0.0103            & 0.01             & 0.001             & NA \\
L2 Reg (weight\_decay)       & 0                 & 0                & 0                 & NA \\
Dropout                      & 0.1601            & 0.1              & 0.6               & NA \\
LR Decay                     & 0.00417           & 0                & 0                 & NA \\
Batch Norm                   & True              & True             & True              & NA \\
Batch Size                   & 256               & 256              & 50                & NA \\
Epochs                       & 500               & 512              & 100               & NA \\
Early Stopping               & False             & True             & True              & NA \\
mod\_alpha                   & NA                & NA               & 0.2               & NA \\
sigma                        & NA                & NA               & 0.1               & NA \\
cuts                         & NA                & NA               & 300               & NA \\ \bottomrule
\end{tabular}
\caption{Hyperparameters across DeepSurv, CoxTime, and DeepHit models.}
\label{tab:hyperparams_comparison}
\end{table}


```{r, eval=FALSE}

# We could keep the same folds and indices since the dataframes are the same
# Number of folds
#K <- 5 
#n <- nrow(mb)  
# R <- 1
# Interesting times for comparison
all_predictions <- list()
survival_curves <- list()
#numeric_cols <- c("X1", "X2", "X3", "X4","X5", "X6", "X7", "X8", "X9")
numeric_cols <- c("X1", "X2", "X3", "X4", "X9")
#seed <- sample(1:1e6, 1)
#set.seed(seed)


#reticulate::py_run_string("import torch; torch.manual_seed(42)")

reset_torch_seed <- function(seed = 123) {
  reticulate::py_run_string(sprintf("
import torch
import numpy as np
import random
torch.manual_seed(%d)
np.random.seed(%d)
random.seed(%d)
torch.backends.cudnn.deterministic = True
torch.backends.cudnn.benchmark = False
", seed, seed, seed))
}

set.seed(123)
survivalmodels::set_seed(seed_R=123, seed_np = 123, seed_torch = 123) ## important for reproducibility
# If we do not need to preserve the censoring/event proportions per fold iteration
#indices <- sample(seq_len(n))
#folds <- cut(indices, breaks = K, labels = FALSE)

#survival_curves[[r]] <- vector("list", K)

train_idx <- sample(1:n, size = floor(0.8 * n))
mb_train  <- mb[train_idx, ]
mb_test   <- mb[-train_idx,]

t_train_max <- max(mb$time)
# Scale continuous 
means <- sapply(mb_train[, numeric_cols], mean)
sds   <- sapply(mb_train[, numeric_cols], sd)

mb_train[, numeric_cols] <- scale(mb_train[, numeric_cols], center = means, scale = sds)
mb_test[, numeric_cols]  <- scale(mb_test[, numeric_cols], center = means, scale = sds)

#t_train_max <- max(mb_train$time)
mb_train$time <- mb_train$time / t_train_max
mb_test$time <- mb_test$time / t_train_max

ts <- sort(unique(mb_test$time))
# Extract only the covariates used for predictions
test_covariates <- mb_test[, c("X1", "X2", "X3", "X4",
                               "X5", "X6", "X7", "X8", "X9")]
# intervals for prediction with scaling
ts_scaled <- seq(0, round(t_train_max), 1)
approx_seq <- ts_scaled / t_train_max

# Fit models
#cat("Dataset:", i, ".", j, '\n')
#cat("Seed: ", seed, '\n')
#cat("Repetition ", r, '\n')
cat('Train set dimensions:', dim(mb_train), '\n')
cat('Train set event:', mean(mb_train$status == 1)*100, '\n')
cat('Train set censoring:', mean(mb_train$status == 0)*100, '\n')
cat('Test set dimensions:', dim(mb_test), '\n')
cat('Train set event:', mean(mb_test$status == 1)*100, '\n')
cat('Test set censoring:', mean(mb_test$status == 0)*100, '\n')
cat('\n')


# 1) Random Forest:
# Fit the model
rf <- randomForestSRC::rfsrc(Surv(time, status) ~ X1 + X2 +
                               X3 + X4 + X5 + X6 + 
                               X7 + X8 + X9, 
                             data = mb_train, ntree = 100)
# Predict survival object
surv_rf_object <- predict(rf, newdata = mb_test, type = "surv")
surv_rf <- surv_rf_object$survival
time_grid <- surv_rf_object$time.interest# * t_train_max
# Interpolate to have same times in all survival curves
surv_rf_int <- apply(surv_rf, 1, function(s){
  approx(x = time_grid, y=s, xout = approx_seq, rule = 2)$y
})
surv_rf_int <- t(surv_rf_int)
rownames(surv_rf_int) <- rownames(mb_test)
colnames(surv_rf_int) <- ts_scaled
# Calculate expected mortality
rf_exp_mort <- rowSums(-log(pmax(surv_rf_int, 1e-10)))


# 2) Predict risk Cox PH
# Fit with cox ph 
cox_ph <- survival::coxph(Surv(time, status) ~ X1 + X2 + 
                           X3 + X4 + X5 + X6 + 
                           X7 + X8 + X9, 
                             data = mb_train)

# Predict survival object
sf <- survfit(cox_ph, newdata = mb_test)
surv_cox <- t(sf$surv)
time_grid <- sf$time #* t_train_max
#Interpolate
surv_cox_int <- apply(surv_cox, 1, function(s){
   approx(x = time_grid, y=s, xout = approx_seq, rule = 2)$y
})

surv_cox_int <- t(surv_cox_int)
rownames(surv_cox_int) <- rownames(mb_test)
colnames(surv_cox_int)  <- ts_scaled
# Calculate expected mortality
cox_exp_mort <- rowSums(-log(pmax(surv_cox_int, 1e-10)))

reset_torch_seed()

# 3) DeepSurv
# Hyperparameters: https://github.com/jaredleekatzman/DeepSurv/blob/41eed003e5b892c81e7855e400861fa7a2d9da4f/experiments/deepsurv/models/metabric_IHC4_clinical_adam_0.json
deepsurv_model <- survivalmodels::deepsurv(Surv(time, status) ~ 
                                             X1 + X2 + 
                                             X3 + X4 + X5 + X6 + 
                                             X7 + X8 + X9, 
                                           data = mb_train, 
                                           frac = 0.2, 
                                           dropout = 0.160087890625,
                                           optimizer = "adam",
                                           activation = "selu", 
                                           batch_norm = TRUE,
                                           num_nodes = c(41), 
                                           #verbose = TRUE,
                                           batch_size = 256L, # ? default, nowhere to be found
                                           learning_rate = 0.010289691253027908, 
                                           lr_decay = 0.0041685546875, 
                                           momentum =  0.8439658203125,# if adam this is not used, is used for sgd
                                           #weight_decay = 1.269e-3, # ? weight_decay > 0 for L2 regularization
                                           #weight_decay = 10.890986328125, # this is too high it cannot be 
                                           epochs = 500) 
# Predict survival object
surv_deepsurv <- predict(deepsurv_model,
                         newdata = mb_test, type = "surv")
time_grid <- as.numeric(colnames(surv_deepsurv))# * t_train_max

# Interpolate
surv_deepsurv_int <- apply(surv_deepsurv, 1, function(s){
  approx(x = time_grid, y=s, xout = approx_seq, rule = 2)$y
})
surv_deepsurv_int <- t(surv_deepsurv_int)
rownames(surv_deepsurv_int) <- rownames(mb_test)
colnames(surv_deepsurv_int) <- ts_scaled
# Calculate expected mortality
deepsurv_exp_mort <- rowSums(-log(pmax(surv_deepsurv_int, 1e-10)))

reset_torch_seed()

# 4) Coxtime
# Hyperparameters from:
# https://github.com/havakv/pycox/blob/3eccdd7fd9844a060f50fdcc315659f33a2d2dc1/examples/cox-time.ipynb
coxtime_model <- survivalmodels::coxtime(Surv(time, status) ~ 
                                           X1 + X2 + X3 + 
                                           X4 + X5 + X6 + 
                                            X7 + X8 + X9, 
                                         data = mb_train, 
                                         optimizer = "adam",
                                         learning_rate = 0.01,       
                                         betas = c(0.9, 0.999), # defaults
                                         activation = "relu",
                                         num_nodes = c(32L, 32L),
                                         batch_norm = TRUE,
                                         dropout = 0.1,
                                         batch_size = 256,
                                         epochs = 512,
                                         early_stopping = TRUE, 
                                         frac = 0.2)

# Predict survival object
surv_coxtime <- predict(coxtime_model,
                        newdata = mb_test, type = "surv")
time_grid <-  as.numeric(colnames(surv_coxtime))# * t_train_max
# Interpolate
surv_coxtime_int <- apply(surv_coxtime, 1, function(s){
  approx(x = time_grid, y=s, xout = approx_seq, rule = 2)$y
})
surv_coxtime_int <- t(surv_coxtime_int)
rownames(surv_coxtime_int) <- rownames(mb_test)
colnames(surv_coxtime_int) <- ts_scaled
# Calculate expected mortality
coxtime_exp_mort <- rowSums(-log(pmax(surv_coxtime_int, 1e-10)))


# scale back the train time too 
mb_train$time <- mb_train$time * t_train_max
mb_test$time <- mb_test$time * t_train_max

reset_torch_seed()

# 5) Deephit
# Original scale for times applied 
# Hyperparameters: https://github.com/havakv/pycox/blob/3eccdd7fd9844a060f50fdcc315659f33a2d2dc1/examples/deephit.ipynb and DeepHit paper, and also double checked with DeSurv paper (some parameters are slightly different)
deephit_model <- survivalmodels::deephit(Surv(time, status) ~ 
                           X1 + X2 + 
                           X3 + X4 + X5 + X6 + 
                           X7 + X8 + X9, 
                           data = mb_train, 
                           optimizer = "adam",
                           activation = "relu", # in DeSurv, ipynb
                           num_nodes = c(32L, 32L), # in ipynb and DeSurv
                           batch_norm = TRUE, # in ipynb
                           dropout = 0.6, #  0.1 in ipynb and 0.6 in DeepHit paper
                           cuts = 300, # in DeSurv and DeepHit, 10 and interpolation in ipynb
                           mod_alpha = 0.2, # in DeSurv, ipynb
                           early_stopping = TRUE, # in DeepHit and ipynb
                           #tolerance = 3,
                           sigma = 0.1, # in DeSurv and ipynb
                           batch_size = 50, # 50 in Deephit, 256 in ipynb
                           epochs = 100, # in ipynb
                           #verbose = TRUE,
                           learning_rate = 0.001, # 0.001 in Deephit, 0.01 in ipynb
                           #betas = c(0.9, 0.999), # default adam
                           frac = 0.2) # validation 
# Survival object
surv_deephit <- predict(deephit_model,
                        newdata = mb_test, type = "surv")
time_grid <- as.numeric(colnames(surv_deephit))# * t_train_max
# Interpolate
surv_deephit_int <- apply(surv_deephit, 1, function(s){
  approx(x = time_grid, y=s, xout = ts_scaled, rule = 2)$y
})
surv_deephit_int <- t(surv_deephit_int)
rownames(surv_deephit_int) <- rownames(mb_test)
colnames(surv_deephit_int) <- ts_scaled
# Calculate expected mortality
deephit_exp_mort <- rowSums(-log(pmax(surv_deephit_int, 1e-10)))


## Gather the survival curves for plotting
survival_curves <- list(patients_ids = rownames(mb_test),
                               test_time = mb_test$time,
                               test_status = mb_test$status,
                               RSF = as.data.frame(surv_rf_int),
                               CoxPH = as.data.frame(surv_cox_int),
                               DeepHit = as.data.frame(surv_deephit_int),
                               DeepSurv = as.data.frame(surv_deepsurv_int),
                               Coxtime = as.data.frame(surv_coxtime_int),
                               covariates = test_covariates)

# Store the results:
fold_results <- data.frame(ho_repeat = "DeepSurvSplits",
                           #cv_fold = k,
                           patients_ids = rownames(mb_test), 
                           test_time = mb_test$time,
                           test_status = mb_test$status,
                           ExpMort.RSF = rf_exp_mort,
                           ExpMort.CoxPH = cox_exp_mort,
                           ExpMort.DeepHit = deephit_exp_mort,
                           ExpMort.DeepSurv = deepsurv_exp_mort,
                           ExpMort.CoxTime = coxtime_exp_mort,
                           RSF = as.data.frame(surv_rf_int),
                           CoxPH = as.data.frame(surv_cox_int),
                           DeepHit = as.data.frame(surv_deephit_int), 
                           DeepSurv = as.data.frame(surv_deepsurv_int),
                           CoxTime = as.data.frame(surv_coxtime_int), 
                           covariates = test_covariates)
 

# Append fold results to the list
all_predictions[[paste("DeepSurvSplits")]] <- fold_results


stacked_predictions <- do.call(rbind, all_predictions)

```


```{r}
#plot_survival_curves(surv_deepsurv, title = "DeepSurv Survival Curves")
```

```{r}

# Plot 
plot_survival_curves(survival_curves$DeepHit, title = "DeepHit Survival Curves")
plot_survival_curves(survival_curves$DeepSurv, title = "DeepSurv Survival Curves")
plot_survival_curves(survival_curves$CoxPH, title = "Cox PH Survival Curves")
plot_survival_curves(survival_curves$Coxtime, title = "Cox Time Survival Curves")
plot_survival_curves(survival_curves$RSF, title = "RSF Survival Curves")
```


Bootstrap 100 times the test set that is stacked_predictions, the size of the bootstrapped sample is equal to the test set. 

## Study 1:

```{r}
set.seed(123)

n_bootstraps <- 100

dataset_size <- nrow(stacked_predictions) 

# re sample by patient id
resample_indices <- replicate(n_bootstraps, 
                              sample(stacked_predictions$patients_ids, 
                                     size = dataset_size, replace = TRUE), 
                              simplify = FALSE)
```

```{r}
subset_expm <- list()
subset_surv <- list()

for (r in seq_along(resample_indices)) {
  subset_expm[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions, 
                 model_names = "all",
                 input_type = "ExpectedMortality",
                 bootstrap_patient_ids = resample_indices[[r]])
  subset_surv[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions, 
                model_names = "all",
                input_type = "Distribution",
                bootstrap_patient_ids = resample_indices[[r]])
  
  
}

```



```{r}
select_exps <- grep("ExpMort\\.",colnames(subset_expm[[1]]), value = TRUE)
results_expm1 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm1[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("Hmisc::rcorr.cens",
                                                        "survC1::Est.Cval", 
                                                        "SurvMetrics::Cindex", 
                                                        "lifelines", 
                                                        "sksurv.censored"),
                                  eval.times = round(max(mb$time)), #C
                                  sampled_data = subset_expm)
}

select_exps <- grep("ExpMort\\.",colnames(subset_expm[[1]]), value = TRUE)
results_expm2 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm2[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pysurvival",
                                                        "pec::cindex",
                                                        "survival.n", 
                                                        "survival.n/G2",
                                                        "sksurv.ipcw"),
                                  eval.times = round(max(mb$time[mb$status == 1])), # C tau.
                                  sampled_data = subset_expm, 
                                  additional=list(
                              sksurv_train_time = mb_train$time, 
                              sksurv_train_status = mb_train$status, 
                              sksurv_tied_tol = 1e-8))
}


model_names <- unique(sub("\\..*", "", grep("^[A-Za-z]+\\.\\d+$",
                                            colnames(subset_surv[[1]]), value = TRUE)))
results_surv <- list()
for (model in model_names) {
  # Antolinis
  cat("Calculating bootstrap for Distribution for Model = ", model, "\n")
  results_surv[[model]] <-  bootstrap.metric(metrics.wrapper,
                                  dataset=list(
                                     predicted = model,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pycox.Ant", "pycox.Adj.Ant"),
                                  sampled_data = subset_surv)
}
results_expm3 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm3[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("survC1::Est.Cval",
                                                        "pec::cindex",
                                                        "survival.n", 
                                                        "survival.n/G2"),
                                  eval.times = 120, #C
                                  sampled_data = subset_expm)
}
```



```{r, fig.width=13, fig.height=6}
# Extract the plots
expm_df_plot1 <- make_expm_plot_entries(results_expm1)
expm_df_plot2 <- make_expm_plot_entries(results_expm2)
expm_df_plot3 <- make_expm_plot_entries(results_expm3)


expm_df_plot1$Notation <- "C"
expm_df_plot2$Notation <- "C_tau"
expm_df_plot3$Notation <- "C_tau_10"  # if needed
surv_df_plot$Notation  <- "C_td"  # or whatever makes sense

df_plot <- bind_rows(expm_df_plot1, expm_df_plot2, expm_df_plot3, surv_df_plot)
df_plot$Metric <- sub("::.*", "", df_plot$Metric)

df_plot$Notation <- factor(df_plot$Notation,
  levels = c("C", "C_tau_10", "C_tau", "C_td"),
  labels = c(
    "tilde(C)~(Expected~Mortality)",
    "tilde(C)[tau]~(Expected~Mortality~tau==10~years)",
    "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
    "tilde(C)[td]~(Survival~Distribution)"
  )
)

df_plot$Model <- factor(df_plot$Model, levels = c("DeepSurv", "CoxTime", "CoxPH", "RSF", "DeepHit"))

df_plot$Metric <- factor(df_plot$Metric, levels = unique(df_plot$Metric))
# Get numeric positions of each Metric for vertical lines
metric_levels <- levels(df_plot$Metric)
n_metrics <- length(metric_levels)

# Define where to put the vertical lines between metrics
separator_positions <- seq(1.5, n_metrics - 0.5, by = 1)

### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.4), size = 0.5) +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
   geom_vline(xintercept = separator_positions, 
              linetype = "dashed", color = "grey70") +
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank())
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.5, color = "black") +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   geom_vline(xintercept = separator_positions, 
              linetype = "dashed", color = "grey70") +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank())

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```










### Study 3:

Trying different implementations:

```{r}
a <- make_expm_table(results_expm1)
b <- make_expm_table(results_expm2)
b$InputType <- "Exp.Mort tau=10y"
c <- make_expm_table(results_expm3)
c$InputType <- "Exp.Mort tau=max(T, delta=1)"
d <- make_surv_table(results_surv)


```




### Table with results

Gather all the results


```{r}
df_combined <- bind_rows(a, b, c, d)

notation_map <- data.frame(
  Metric = c("pycox.Ant", "pycox.Adj.Ant", 
             "pec", "survC1", "survival.n/G2", "sksurv.ipcw","survival.n", 
             "Hmisc", "SurvMetrics", "lifelines", "pysurvival", 
             "sksurv.censored"),
  Notation = c("$C_{td}$", "$C_{td}$", 
               "$C_{\\tau}$", "$C_{\\tau}$", "$C_{\\tau}$", 
               "$C_{\\tau}$", "$C_{\\tau}$",
               "$C$", "$C$", "$C$", "$C$", "$C$"),
  stringsAsFactors = FALSE
)

df_combined$Metric <- sub("::.*", "", df_combined$Metric)
df_combined <- merge(df_combined, notation_map, by= "Metric", all.x = TRUE)

```


#### Save table

```{r}
library(kableExtra)

df_combined_bolded <- df_combined

for (i in 1:nrow(df_combined_bolded)) {
  row <- df_combined_bolded[i, model_names]
  # only the mean, not the ci
  numeric_values <- as.numeric(sub(" .*", "", row))

  max_index <- which.max(numeric_values)
  
  # only bold the mean
  row[max_index] <- sub("^(\\d+\\.\\d+)", "\\\\textbf{\\1}", row[max_index])
  
  df_combined_bolded[i, model_names] <- row
}

df_combined_bolded <- df_combined_bolded %>% arrange(factor(InputType, levels = c("Distrib", "Exp.Mort", "Exp.Mort tau=10y", "Exp.Mort tau=max(T, delta=1)" )))


latex_code <- kable(df_combined_bolded, format = "latex", 
                    booktabs = TRUE, 
                    escape = FALSE)

writeLines(latex_code, "Paper/Tables/results_tableHoldOut.tex")
#writeLines(latex_code, "Results/PaperTables/results_table5fold_Seed123.tex")

```

