---
title: "MetabricPointEstim2"
output: html_document
date: "2025-04-15"
---

## Concordance index multiverse.


Load libraries

```{r, include=FALSE}
# Survival metrics
library(reticulate)
library(arrow)
library(caret)
library(riskRegression)
library(prodlim)
library(pec)
library(survival)
library(rhdf5)
library(randomForestSRC)
library(survAUC)
library(Hmisc)
library(dplyr)
# Plotting
library(gridExtra)
# Parallelization
library(doFuture)
library(future)
library(progressr)
library(foreach)
library(MASS)
library(flexsurv)
library(furrr)
library(pysurvivalR)
library(survivalmodels)

library(tidyr)
```


```{r setup, include=FALSE}
#Sys.unsetenv("RETICULATE_PYTHON")
Sys.setenv(OMP_NUM_THREADS = "1")       # Limits OpenMP to 1 thread
Sys.setenv(NUMBA_NUM_THREADS = "1")     # Limits Numba to 1 thread
Sys.setenv(MKL_NUM_THREADS = "1")       # Limits Intel MKL to 1 thread
Sys.setenv(KMP_WARNINGS = "0")          # Disables OpenMP warnings
Sys.setenv(OPENBLAS_NUM_THREADS = "1")  # Limits OpenBLAS to 1 thread

# #Sys.setenv(NUMBA_DISABLE_JIT = "0")  
# library(reticulate)

#use_condaenv("/opt/homebrew/Caskroom/miniforge/base/envs/py-rstudio", required=TRUE)
#use_virtualenv("~/.virtualenvs/venv-DeSurv_python_R")
#use_python("/opt/homebrew/Caskroom/miniforge/base/envs/venv-DeSurv3/bin/python3.9")
#py_config()
```


```{r}
# read the processed file
mb <- read.table("./Datasets/metabric_preprocess_multiverse.csv", sep="," , header = TRUE)

# Keep the id as index
rownames(mb) <- mb$PATIENT_ID

# Remove the id column
mb <- mb[,2:12]

# Sote the previous colnames to keep track of them
cols <- colnames(mb)

# Asign new column names
newcols <- c("X1", "X2", "X3", "X4","X5", "X6", "X7", "X8", "X9", "time", "status")
colnames(mb) <- newcols

# NO NEED OF SUMING, there is no 0s... 
min(mb$time)
#mb$time <- mb$time + 1e-8

# Order according to time and status
mb[order(mb$time, -mb$status),]

```
```{r}
source("./CindexHelperFunctions.R")
```



```{r}
head(mb)
```
```{r}
cat("Number of patients: ", nrow(mb), "\n")
cat("Number of unique times: ", length(unique(mb$time)), "\n")

hist(mb$time)


table_times <- table(mb$time)
ties_times <- table_times[table_times > 1]

length(ties_times)

num_tied_pairs <- sum(choose(ties_times, 2))
n_tied_pairs <- sum((ties_times * (ties_times - 1)) / 2)
cat("Number of pairs that have tied times: ", num_tied_pairs, "\n")
```

#### CROSSVALIDATION ONCE 

Predict on the k-test fold out of 5 folds with different models

For the hyperparameters of deep learning models that have been trained with survivalmodels package is relevant to look at the pytorch optimizer: https://raphaels1.github.io/survivalmodels/reference/get_pycox_optim.html 

Also the parameters for each model:
1) Deephit: https://raphaels1.github.io/survivalmodels/reference/deephit.html
2) DeepSurv: https://raphaels1.github.io/survivalmodels/reference/deepsurv.html
3) CoxTime: https://raphaels1.github.io/survivalmodels/reference/coxtime.html

\begin{table}[ht]
\centering
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Hyper-parameter}     & \textbf{DeepSurv} & \textbf{CoxTime} & \textbf{DeepHit}  & RandomForest \\ \midrule
Optimizer                    & adam              & adam             & adam              & NA \\
Activation                   & selu              & relu             & relu              & NA\\
\# Dense Layers              & 1                 & 2                & 2                 & - \\
\# Nodes / Layer             & 41                & 32, 32           & 32, 32            & 100 trees\\
Learning Rate                & 0.0103            & 0.01             & 0.001             & NA \\
L2 Reg (weight\_decay)       & 0                 & 0                & 0                 & NA \\
Dropout                      & 0.1601            & 0.1              & 0.6               & NA \\
LR Decay                     & 0.00417           & 0                & 0                 & NA \\
Batch Norm                   & True              & True             & True              & NA \\
Batch Size                   & 256               & 256              & 50                & NA \\
Epochs                       & 500               & 512              & 100               & NA \\
Early Stopping               & False             & True             & True              & NA \\
mod\_alpha                   & NA                & NA               & 0.2               & NA \\
sigma                        & NA                & NA               & 0.1               & NA \\
cuts                         & NA                & NA               & 300               & NA \\ \bottomrule
\end{tabular}
\caption{Hyperparameters across DeepSurv, CoxTime, and DeepHit models.}
\label{tab:hyperparams_comparison}
\end{table}

```{r, eval=FALSE}

## Reproducibility
reset_torch_seed <- function(seed = 123) {
  reticulate::py_run_string(sprintf("
import torch
import numpy as np
import random
torch.manual_seed(%d)
np.random.seed(%d)
random.seed(%d)
torch.backends.cudnn.deterministic = True
torch.backends.cudnn.benchmark = False
", seed, seed, seed))
}

set.seed(123)
survivalmodels::set_seed(seed_R=123, seed_np = 123, seed_torch = 123) 

### Crossval
# Number of folds
K <- 5 
n <- nrow(mb)  

# Interesting times for comparison
#ts <- sort(unique(round(mb$time)))
all_predictions <- list()

# List of numeric column for standarization
numeric_cols <- c("X1", "X2", "X3", "X4", "X9")

# Create empty list for survival curves
survival_curves <- vector("list", K)

# Shuffle indices and create folds
#indices <- sample(seq_len(n))
#folds <- cut(indices, breaks = K, labels = FALSE)
folds <- createFolds(mb$status, k = K, list = TRUE) # return test sets

# Interesting intervals for prediction
mb_t_max <- round(max(mb$time))
ts_scaled <- seq(0, mb_t_max, 1)

for (k in 1:K) {

  # Define test and training indices
  #mb_test_idx  <- which(folds == k)
  #mb_train_idx <- setdiff(seq_len(n), mb_test_idx)
  
  # Define test and training indices
  mb_test_idx  <- folds[[k]]
  mb_train_idx <- setdiff(seq_len(nrow(mb)), mb_test_idx)
  
  # Subset the dataset
  mb_train <- mb[mb_train_idx, ]
  mb_test  <- mb[mb_test_idx, ]
  
  t_train_max <- max(mb_train$time)
  
  # Scale continuous 
  means <- sapply(mb_train[, numeric_cols], mean)
  sds   <- sapply(mb_train[, numeric_cols], sd)

  mb_train[, numeric_cols] <- scale(mb_train[, numeric_cols],
                                    center = means, scale = sds)
  mb_test[, numeric_cols]  <- scale(mb_test[, numeric_cols],
                                    center = means, scale = sds)

  #t_train_max <- max(mb_train$time)
  mb_train$time <- mb_train$time / t_train_max
  mb_test$time <- mb_test$time / t_train_max
  
  # Extract only the covariates used for predictions
  test_covariates <- mb_test[, c("X1", "X2", "X3", "X4",
                                 "X5", "X6", "X7", "X8", "X9")]
  # intervals for prediction scaled
  approx_seq <- ts_scaled / t_train_max
  # Fit models
  #cat("Dataset:", i, ".", j, '\n')
  cat("Fold:", k, '\n')
  cat("tmax:", t_train_max, "\n")
  cat('Train set dimensions:', dim(mb_train), '\n')
  cat('Train set event:', mean(mb_train$status == 1)*100, '\n')
  cat('Train set censoring:', mean(mb_train$status == 0)*100, '\n')
  cat('Test set dimensions:', dim(mb_test), '\n')
  cat('Test set event:', mean(mb_test$status == 1)*100, '\n')
  cat('Test set censoring:', mean(mb_test$status == 0)*100, '\n')
  cat('\n')
  
  
  # 1) Random Forest:
  # Fit the model
  rf <- randomForestSRC::rfsrc(Surv(time, status) ~ X1 + X2 +
                                 X3 + X4 + X5 + X6 + 
                                 X7 + X8 + X9, 
                               data = mb_train, ntree = 100)
  # Predict survival object
  surv_rf_object <- predict(rf, newdata = mb_test, type = "surv")
  surv_rf <- surv_rf_object$survival
  time_grid <- surv_rf_object$time.interest# * t_train_max
  # Interpolate to have same times in all survival curves
  surv_rf_int <- apply(surv_rf, 1, function(s){
    approx(x = time_grid, y=s, xout = approx_seq, rule = 2)$y
  })
  surv_rf_int <- t(surv_rf_int)
  rownames(surv_rf_int) <- rownames(mb_test)
  colnames(surv_rf_int) <- ts_scaled
  # Calculate expected mortality
  rf_exp_mort <- rowSums(-log(pmax(surv_rf_int, 1e-10)))

  
  # 2) Predict risk Cox PH
  # Fit with cox ph 
  cox_ph <- survival::coxph(Surv(time, status) ~ X1 + X2 + 
                             X3 + X4 + X5 + X6 + 
                             X7 + X8 + X9, 
                               data = mb_train)
  
  # Predict survival object
  sf <- survfit(cox_ph, newdata = mb_test)
  surv_cox <- t(sf$surv)
  time_grid <- sf$time #* t_train_max
  #Interpolate
  surv_cox_int <- apply(surv_cox, 1, function(s){
     approx(x = time_grid, y=s, xout = approx_seq, rule = 2)$y
  })
  
  surv_cox_int <- t(surv_cox_int)
  rownames(surv_cox_int) <- rownames(mb_test)
  colnames(surv_cox_int)  <- ts_scaled
  # Calculate expected mortality
  cox_exp_mort <- rowSums(-log(pmax(surv_cox_int, 1e-10)))

  reset_torch_seed()
  
  # 3) DeepSurv
  # Hyperparameters: https://github.com/jaredleekatzman/DeepSurv/blob/41eed003e5b892c81e7855e400861fa7a2d9da4f/experiments/deepsurv/models/metabric_IHC4_clinical_adam_0.json
  deepsurv_model <- survivalmodels::deepsurv(Surv(time, status) ~ 
                                               X1 + X2 + 
                                               X3 + X4 + X5 + X6 + 
                                               X7 + X8 + X9, 
                                             data = mb_train, 
                                             frac = 0.2, 
                                             dropout = 0.160087890625,
                                             optimizer = "adam",
                                             activation = "selu", 
                                             batch_norm = TRUE,
                                             num_nodes = c(41), 
                                             #verbose = TRUE,
                                             batch_size = 256L, # ? default, nowhere to be found
                                             learning_rate = 0.010289691253027908, 
                                             lr_decay = 0.0041685546875, 
                                             momentum =  0.8439658203125,# if adam this is not used, is used for sgd
                                             #weight_decay = 1.269e-3, # ? weight_decay > 0 for L2 regularization
                                             #weight_decay = 10.890986328125, # this is too high it cannot be 
                                             epochs = 500) 
  # Predict survival object
  surv_deepsurv <- predict(deepsurv_model,
                           newdata = mb_test, type = "surv")
  time_grid <- as.numeric(colnames(surv_deepsurv))# * t_train_max
  
  # Interpolate
  surv_deepsurv_int <- apply(surv_deepsurv, 1, function(s){
    approx(x = time_grid, y=s, xout = approx_seq, rule = 2)$y
  })
  surv_deepsurv_int <- t(surv_deepsurv_int)
  rownames(surv_deepsurv_int) <- rownames(mb_test)
  colnames(surv_deepsurv_int) <- ts_scaled
  # Calculate expected mortality
  deepsurv_exp_mort <- rowSums(-log(pmax(surv_deepsurv_int, 1e-10)))
  
  reset_torch_seed()
  
  # 4) Coxtime
  # Hyperparameters from:
  # https://github.com/havakv/pycox/blob/3eccdd7fd9844a060f50fdcc315659f33a2d2dc1/examples/cox-time.ipynb
  coxtime_model <- survivalmodels::coxtime(Surv(time, status) ~ 
                                             X1 + X2 + X3 + 
                                             X4 + X5 + X6 + 
                                              X7 + X8 + X9, 
                                           data = mb_train, 
                                           optimizer = "adam",
                                           learning_rate = 0.01,       
                                           betas = c(0.9, 0.999), # defaults
                                           activation = "relu",
                                           num_nodes = c(32L, 32L),
                                           batch_norm = TRUE,
                                           dropout = 0.1,
                                           batch_size = 256,
                                           epochs = 512,
                                           early_stopping = TRUE, 
                                           frac = 0.2)
  
  # Predict survival object
  surv_coxtime <- predict(coxtime_model,
                          newdata = mb_test, type = "surv")
  time_grid <-  as.numeric(colnames(surv_coxtime))# * t_train_max
  # Interpolate
  surv_coxtime_int <- apply(surv_coxtime, 1, function(s){
    approx(x = time_grid, y=s, xout = approx_seq, rule = 2)$y
  })
  surv_coxtime_int <- t(surv_coxtime_int)
  rownames(surv_coxtime_int) <- rownames(mb_test)
  colnames(surv_coxtime_int) <- ts_scaled
  # Calculate expected mortality
  coxtime_exp_mort <- rowSums(-log(pmax(surv_coxtime_int, 1e-10)))
  
  
  # scale back the train time too 
  mb_train$time <- mb_train$time * t_train_max
  mb_test$time <- mb_test$time * t_train_max
  
  reset_torch_seed()
  
  # 5) Deephit
  # Original scale for times applied 
  # Hyperparameters: https://github.com/havakv/pycox/blob/3eccdd7fd9844a060f50fdcc315659f33a2d2dc1/examples/deephit.ipynb and DeepHit paper, and also double checked with DeSurv paper (some parameters are slightly different)
  deephit_model <- survivalmodels::deephit(Surv(time, status) ~ 
                             X1 + X2 + 
                             X3 + X4 + X5 + X6 + 
                             X7 + X8 + X9, 
                             data = mb_train, 
                             optimizer = "adam",
                             activation = "relu", # in DeSurv, ipynb
                             num_nodes = c(32L, 32L), # in ipynb and DeSurv
                             batch_norm = TRUE, # in ipynb
                             dropout = 0.6, #  0.1 in ipynb and 0.6 in DeepHit paper
                             cuts = 300, # in DeSurv and DeepHit, 10 and interpolation in ipynb
                             mod_alpha = 0.2, # in DeSurv, ipynb
                             early_stopping = TRUE, # in DeepHit and ipynb
                             #tolerance = 3,
                             sigma = 0.1, # in DeSurv and ipynb
                             batch_size = 50, # 50 in Deephit, 256 in ipynb
                             epochs = 100, # in ipynb
                             #verbose = TRUE,
                             learning_rate = 0.001, # 0.001 in Deephit, 0.01 in ipynb
                             #betas = c(0.9, 0.999), # default adam
                             frac = 0.2) # validation 

  # Survival object
  surv_deephit <- predict(deephit_model,
                          newdata = mb_test, type = "surv")
  time_grid <- as.numeric(colnames(surv_deephit))# * t_train_max
  # Interpolate
  surv_deephit_int <- apply(surv_deephit, 1, function(s){
    approx(x = time_grid, y=s, xout = ts_scaled, rule = 2)$y
  })
  surv_deephit_int <- t(surv_deephit_int)
  rownames(surv_deephit_int) <- rownames(mb_test)
  colnames(surv_deephit_int) <- ts_scaled
  # Calculate expected mortality
  deephit_exp_mort <- rowSums(-log(pmax(surv_deephit_int, 1e-10)))
  

  # Gather the survival survs too
  survival_curves[[k]] <- list(patients_ids = rownames(mb_test),
                               test_time = mb_test$time,
                               test_status = mb_test$status,
                               RSF = as.data.frame(surv_rf_int),
                               CoxPH = as.data.frame(surv_cox_int),
                               DeepHit = as.data.frame(surv_deephit_int),
                               DeepSurv = as.data.frame(surv_deepsurv_int),
                               Coxtime = as.data.frame(surv_coxtime_int),
                               covariates = test_covariates)
  
  fold_results <- data.frame(cv_fold = k,
                             patients_ids = rownames(mb_test), 
                             test_time = mb_test$time,
                             test_status = mb_test$status,
                             ExpMort.RSF = rf_exp_mort,
                             ExpMort.CoxPH = cox_exp_mort,
                             ExpMort.DeepHit = deephit_exp_mort,
                             ExpMort.DeepSurv = deepsurv_exp_mort,
                             ExpMort.CoxTime = coxtime_exp_mort,
                             RSF = as.data.frame(surv_rf_int),
                             CoxPH = as.data.frame(surv_cox_int),
                             DeepHit = as.data.frame(surv_deephit_int), 
                             DeepSurv = as.data.frame(surv_deepsurv_int),
                             CoxTime = as.data.frame(surv_coxtime_int), 
                             covariates = test_covariates)
    

  # Append fold results to the list
 all_predictions[[paste("Fold", k)]] <- fold_results

}

stacked_predictions <- do.call(rbind, all_predictions)

```


```{r}
stacked_predictions <- readRDS("./Results/5foldCV_MetabricStackedPredictions.rds")
survival_curves     <- readRDS("./Results/5foldCV_MetabricSurvivalCurves.rds")
```

```{r}
source("./CindexHelperFunctions.R")
```

```{r}
# Plot 
plot_survival_curves(survival_curves[[1]]$DeepHit, title = "DeepHit Survival Curves")
plot_survival_curves(survival_curves[[1]]$DeepSurv, title = "DeepSurv Survival Curves")
plot_survival_curves(survival_curves[[1]]$CoxPH, title = "Cox PH Survival Curves")
plot_survival_curves(survival_curves[[1]]$Coxtime, title = "Cox Time Survival Curves")
plot_survival_curves(survival_curves[[1]]$RSF, title = "RSF Survival Curves")
```



Need to be able to calculate the C-index in 3 different ways: 

1) Comparing probabilities at a specific time point. What Le Blanche does not like. 
And a summary distribution like median times...

2) With Expected mortality transformation

3) With the survival distribution. 


Create bootstrap samples of 1000 patients each
```{r, eval=FALSE}
# TEST

set.seed(123)

#subset_stacked <- stacked_predictions[(stacked_predictions$cv_repeat == 1),]

# Number of bootstrapps
n_bootstraps <- 100
# Size of boostrapped sample
size_sample <- 1000

# re sample by patient id
resample_indices <- replicate(n_bootstraps, 
                              sample(stacked_predictions$patients_ids, 
                                     size = size_sample, replace = TRUE), 
                              simplify = FALSE)
```


### TIES IN TIMES

```{r}
# total possible pairs in Metabric
n_total_pairs <- ((nrow(stacked_predictions))*(nrow(stacked_predictions)-1)) / 2

```

```{r}
cat("Number of patients: ", nrow(mb), "\n")
cat("Number of unique times: ", length(unique(mb$time)), "\n")

# ggplot(mb, aes(x = time)) +
#   geom_histogram(breaks = seq(0, max(mb$time, na.rm = TRUE), by = 3), 
#                  color = "black", fill = "skyblue") 


table_times <- table(mb$time)
ties_times <- table_times[table_times > 1]

num_tied_pairs <- sum(choose(ties_times, 2))
n_tied_pairs <- sum((ties_times * (ties_times - 1)) / 2)
cat("Number of pairs that have tied times: ", num_tied_pairs, "\n")

cat("% of tied pairs out of total possible pairs ", num_tied_pairs/n_total_pairs *100, "\n")
```

```{r}
cat("New number of unique times: ", length(unique(stacked_predictions$test_time)), "\n")

table_times <- table(stacked_predictions$test_time)
ties_times <- table_times[table_times > 1]

num_tied_pairs <- sum(choose(ties_times, 2))
n_tied_pairs <- sum((ties_times * (ties_times - 1)) / 2)
cat("Number of pairs that have tied times: ", num_tied_pairs, "\n")
cat("% of tied pairs out of total possible pairs ", num_tied_pairs/n_total_pairs *100, "\n")

```

```{r}
stacked_predictions_TT <- stacked_predictions
stacked_predictions_TT$test_time <- round(stacked_predictions_TT$test_time, 1)

cat("New number of unique times: ", length(unique(stacked_predictions_TT$test_time)), "\n")

table_times <- table(stacked_predictions_TT$test_time)
ties_times <- table_times[table_times > 1]

num_tied_pairs <- sum(choose(ties_times, 2))
n_tied_pairs <- sum((ties_times * (ties_times - 1)) / 2)
cat("Number of pairs that have tied times: ", num_tied_pairs, "\n")
cat("% of tied pairs out of total possible pairs ", num_tied_pairs/n_total_pairs *100, "\n")

```


```{r}
stacked_predictions_T2 <- stacked_predictions
stacked_predictions_T2$test_time <- round(stacked_predictions_T2$test_time)

cat("New number of unique times: ", length(unique(stacked_predictions_T2$test_time)), "\n")

table_times <- table(stacked_predictions_T2$test_time)
ties_times <- table_times[table_times > 1]

num_tied_pairs <- sum(choose(ties_times, 2))
n_tied_pairs <- sum((ties_times * (ties_times - 1)) / 2)
cat("Number of pairs that have tied times: ", num_tied_pairs, "\n")
cat("% of tied pairs out of total possible pairs ", num_tied_pairs/n_total_pairs *100, "\n")

```

```{r}
# Bin test_time into 3-month intervals
stacked_predictions_B <- stacked_predictions
# Get bin midpoints
bin_midpoints <- seq(0, max(stacked_predictions_B$test_time, na.rm = TRUE) + 3, by = 3) + 1.5
bin_midpoints <- bin_midpoints[-length(bin_midpoints)]  # remove last as it’s beyond max
stacked_predictions_B$test_time <- cut(stacked_predictions_B$test_time, 
                      breaks = seq(0, max(stacked_predictions_B$test_time, na.rm = TRUE) + 3, by = 3), 
                      right = FALSE, include.lowest = TRUE)

cat("New number of unique times: ", length(unique(stacked_predictions_B$test_time)), "\n")

table_times <- table(stacked_predictions_B$test_time)
ties_times <- table_times[table_times > 1]

num_tied_pairs <- sum(choose(ties_times, 2))
cat("Number of pairs that have tied times: ", num_tied_pairs, "\n")

# Map levels to midpoints
stacked_predictions_B$test_time<- bin_midpoints[as.numeric(stacked_predictions_B$test_time)]


cat("% of tied pairs out of total possible pairs ", num_tied_pairs/n_total_pairs *100, "\n")

```


```{r}
# ties_results_surv <- list(
#   results_surv = results_surv, # no rounding
#   results_surv_B = results_surv_B, # binning
#   results_surv_TT = results_surv_TT, # ties in times 1 
#   results_surv_T2 = results_surv_T2 # times in times 2
# )
```

Try to see how the change RiskAtT for several Ts affects the C-index. Do evaluation at different time points.

For that we can do Harrels to have sth that does not change with time, then the ones that require tau.


```{r}
# subset_risk_10 <- list()
subset_expm <- list()
subset_surv <- list()

for (r in seq_along(resample_indices)) {
  # subset_risk_10[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions, 
  #                model_names = "all",
  #                input_type = "RiskAtT",
  #                specific_time = c(60), # 5 years
  #                bootstrap_patient_ids = resample_indices[[r]])
  subset_expm[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions, 
                 model_names = "all",
                 input_type = "ExpectedMortality",
                 bootstrap_patient_ids = resample_indices[[r]])
  subset_surv[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions, 
                model_names = "all",
                input_type = "Distribution",
                bootstrap_patient_ids = resample_indices[[r]])
  
  
}
```

```{r}
# subset_risk_10_TT <- list()
subset_expm_TT <- list()
subset_surv_TT <- list()

for (r in seq_along(resample_indices)) {
  # subset_risk_10_TT[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TT, 
  #                model_names = "all",
  #                input_type = "RiskAtT",
  #                specific_time = c(60), # 5 years
  #                bootstrap_patient_ids = resample_indices[[r]])
  subset_expm_TT[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TT, 
                 model_names = "all",
                 input_type = "ExpectedMortality",
                 bootstrap_patient_ids = resample_indices[[r]])
  subset_surv_TT[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TT, 
                model_names = "all",
                input_type = "Distribution",
                bootstrap_patient_ids = resample_indices[[r]])
  
  
}
```

```{r}
# subset_risk_10_T2 <- list()
subset_expm_T2 <- list()
subset_surv_T2 <- list()

for (r in seq_along(resample_indices)) {
  # subset_risk_10_T2[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_T2, 
  #                model_names = "all",
  #                input_type = "RiskAtT",
  #                specific_time = c(60), # 5 years
  #                bootstrap_patient_ids = resample_indices[[r]])
  subset_expm_T2[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_T2, 
                 model_names = "all",
                 input_type = "ExpectedMortality",
                 bootstrap_patient_ids = resample_indices[[r]])
  subset_surv_T2[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_T2, 
                model_names = "all",
                input_type = "Distribution",
                bootstrap_patient_ids = resample_indices[[r]])
  
  
}
```

```{r}
# subset_risk_10 <- list()
subset_expm_B <- list()
subset_surv_B <- list()

for (r in seq_along(resample_indices)) {
  # subset_risk_10[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions, 
  #                model_names = "all",
  #                input_type = "RiskAtT",
  #                specific_time = c(60), # 5 years
  #                bootstrap_patient_ids = resample_indices[[r]])
  subset_expm_B[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_B, 
                 model_names = "all",
                 input_type = "ExpectedMortality",
                 bootstrap_patient_ids = resample_indices[[r]])
  subset_surv_B[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_B, 
                model_names = "all",
                input_type = "Distribution",
                bootstrap_patient_ids = resample_indices[[r]])
  
  
}
```


```{r}
# selected_models <- colnames(subset_risk_10[[1]])[4:length(colnames(subset_risk_10[[1]]))]
# 
# results_risk1 <- list()
# for (model in selected_models) {
#    t <- as.numeric(sub(".*\\.", "", model))
#    cat("Calculating bootstrap at eval.time = ", t, "for Model = ", model, "\n")
#    results_risk1[[model]] <- bootstrap.metric.parallel(metrics.wrapper,
#                                   dataset=list(
#                                      predicted = model,
#                                      censoring = "test_status",
#                                      time = "test_time"),
#                                   implementation = list("Hmisc::rcorr.cens",
#                                                         "pysurvival",
#                                                         "SurvMetrics::Cindex", 
#                                                         "lifelines", 
#                                                         "sksurv.censored"),
#                                   eval.times = round(max(mb$time[mb$status == 1])), # or set t to sth else
#                                   sampled_data = subset_risk_10)
# 
# }
# 
# results_risk2 <- list()
# for (model in selected_models) {
#    t <- as.numeric(sub(".*\\.", "", model))
#    cat("Calculating bootstrap at eval.time = ", t, "for Model = ", model, "\n")
#    results_risk2[[model]] <- bootstrap.metric.parallel(metrics.wrapper,
#                                   dataset=list(
#                                      predicted = model,
#                                      censoring = "test_status",
#                                      time = "test_time"),
#                                   implementation = list("survC1::Est.Cval", 
#                                                         "pec::cindex",
#                                                         "survival.n", 
#                                                         "survival.n/G2"),
#                                   eval.times = round(max(mb$time[mb$status == 1])), # or set t to sth else
#                                   sampled_data = subset_risk_10)
# 
# }

```

With the normal amount of ties in times

```{r}
select_exps <- grep("ExpMort\\.",colnames(subset_expm[[1]]), value = TRUE)
results_expm1 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm1[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("Hmisc::rcorr.cens",
                                                        "pysurvival", 
                                                        "SurvMetrics::Cindex", 
                                                        "lifelines", 
                                                        "sksurv.censored"),
                                  eval.times = round(max(mb$time)), #C
                                  sampled_data = subset_expm)
}

select_exps <- grep("ExpMort\\.",colnames(subset_expm[[1]]), value = TRUE)
results_expm2 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm2[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pec::cindex",
                                                        "survC1::Est.Cval", 
                                                        "survival.n", 
                                                        "survival.n/G2"),
                                  eval.times = round(max(mb$time[mb$status == 1])), # C tau.
                                  sampled_data = subset_expm)
}
```

```{r}
model_names <- unique(sub("\\..*", "", grep("^[A-Za-z]+\\.\\d+$",
                                            colnames(subset_surv[[1]]), value = TRUE)))
results_surv <- list()
for (model in model_names) {
  # Antolinis
  cat("Calculating bootstrap for Distribution for Model = ", model, "\n")
  results_surv[[model]] <-  bootstrap.metric(metrics.wrapper,
                                  dataset=list(
                                     predicted = model,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pycox.Adj.Ant", "pycox.Ant"),
                                  sampled_data = subset_surv)
}
```


```{r, fig.width=12, fig.height=6}
expm_df_plot1 <- make_expm_plot_entries(results_expm1)
expm_df_plot2 <- make_expm_plot_entries(results_expm2)
surv_df_plot <- make_surv_plot_entries(results_surv)

df_plot <- bind_rows(expm_df_plot1, expm_df_plot2, surv_df_plot)

notation_map <- data.frame(
  Metric = c("pycox.Ant", "pycox.Adj.Ant", 
             "pec", "survC1", "survival.n/G2", "sksurv.ipcw","survival.n", 
             "Hmisc", "SurvMetrics", "lifelines", "pysurvival", 
             "sksurv.censored"),
  Notation = c("C_td", "C_td", 
               "C_tau", "C_tau", "C_tau", 
               "C_tau", "C_tau",
               "C", "C", "C", "C", "C"),
  stringsAsFactors = FALSE
)
df_plot$Metric <- sub("::.*", "", df_plot$Metric)
df_plot <- merge(df_plot, notation_map, by= "Metric", all.x = TRUE)
df_plot$Notation <- factor(df_plot$Notation, levels = c("C", "C_tau", "C_td"),
                           labels = c(
                             "tilde(C)~(Expected~Mortality)",
                             "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
                             "tilde(C)[td]~(Survival~Distribution)"
                           ))
df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit", "CoxPH", "DeepSurv", "CoxTime"))
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.4, color = "black") +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```



```{r}
select_exps <- grep("ExpMort\\.",colnames(subset_expm_TT[[1]]), value = TRUE)
results_expm1_TT <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm1_TT[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("Hmisc::rcorr.cens",
                                                        "pysurvival",
                                                        "SurvMetrics::Cindex", 
                                                        "lifelines", 
                                                        "sksurv.censored"),
                                  eval.times = round(max(mb$time)), #C
                                  sampled_data = subset_expm_TT)
}

select_exps <- grep("ExpMort\\.",colnames(subset_expm_TT[[1]]), value = TRUE)
results_expm2_TT <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm2_TT[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pec::cindex",
                                                        "survC1::Est.Cval", 
                                                        "survival.n", 
                                                        "survival.n/G2"),
                                  eval.times = round(max(mb$time[mb$status == 1])), # C tau.
                                  sampled_data = subset_expm_TT)
}

model_names <- unique(sub("\\..*", "", grep("^[A-Za-z]+\\.\\d+$",
                                            colnames(subset_surv_TT[[1]]), value = TRUE)))
results_surv_TT <- list()
for (model in model_names) {
  # Antolinis
  cat("Calculating bootstrap for Distribution for Model = ", model, "\n")
  results_surv_TT[[model]] <-  bootstrap.metric(metrics.wrapper,
                                  dataset=list(
                                     predicted = model,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pycox.Ant", "pycox.Adj.Ant"),
                                  sampled_data = subset_surv_TT)
}
```


```{r, fig.width=12, fig.height=6}

expm_df_plot1_TT <- make_expm_plot_entries(results_expm1_TT)
expm_df_plot2_TT <- make_expm_plot_entries(results_expm2_TT)
surv_df_plot_TT <- make_surv_plot_entries(results_surv_TT)

df_plot <- bind_rows(expm_df_plot1_TT, expm_df_plot2_TT, surv_df_plot_TT)

notation_map <- data.frame(
  Metric = c("pycox.Ant", "pycox.Adj.Ant", 
             "pec", "survC1", "survival.n/G2", "sksurv.ipcw","survival.n", 
             "Hmisc", "SurvMetrics", "lifelines", "pysurvival", 
             "sksurv.censored"),
  Notation = c("C_td", "C_td", 
               "C_tau", "C_tau", "C_tau", 
               "C_tau", "C_tau",
               "C", "C", "C", "C", "C"),
  stringsAsFactors = FALSE
)
df_plot$Metric <- sub("::.*", "", df_plot$Metric)
df_plot <- merge(df_plot, notation_map, by= "Metric", all.x = TRUE)
df_plot$Notation <- factor(df_plot$Notation, levels = c("C", "C_tau", "C_td"),
                           labels = c(
                             "tilde(C)~(Expected~Mortality)",
                             "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
                             "tilde(C)[td]~(Survival~Distribution)"
                           ))
df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit", "CoxPH", "DeepSurv", "CoxTime"))
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.4, color = "black") +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```


#### Repeat with more ties


```{r}
select_exps <- grep("ExpMort\\.",colnames(subset_expm_T2[[1]]), value = TRUE)
results_expm1_T2 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm1_T2[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("Hmisc::rcorr.cens",
                                                        "pysurvival", 
                                                        "SurvMetrics::Cindex", 
                                                        "lifelines", 
                                                        "sksurv.censored"),
                                  eval.times = round(max(mb$time)), #C
                                  sampled_data = subset_expm_T2)
}

select_exps <- grep("ExpMort\\.",colnames(subset_expm_T2[[1]]), value = TRUE)
results_expm2_T2 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm2_T2[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pec::cindex",
                                                        "survC1::Est.Cval", 
                                                        "survival.n", 
                                                        "survival.n/G2"),
                                  eval.times = round(max(mb$time[mb$status == 1])), # C tau.
                                  sampled_data = subset_expm_T2)
}
```



```{r}
model_names <- unique(sub("\\..*", "", grep("^[A-Za-z]+\\.\\d+$",
                                            colnames(subset_surv_T2[[1]]), value = TRUE)))
results_surv_T2 <- list()
for (model in model_names) {
  # Antolinis
  cat("Calculating bootstrap for Distribution for Model = ", model, "\n")
  results_surv_T2[[model]] <-  bootstrap.metric(metrics.wrapper,
                                  dataset=list(
                                     predicted = model,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pycox.Ant", "pycox.Adj.Ant"),
                                  sampled_data = subset_surv_T2)
}
```


```{r, fig.width=12, fig.height=6}

expm_df_plot1_T2 <- make_expm_plot_entries(results_expm1_T2)
expm_df_plot2_T2 <- make_expm_plot_entries(results_expm2_T2)
surv_df_plot_T2 <- make_surv_plot_entries(results_surv_T2)

df_plot <- bind_rows(expm_df_plot1_T2, expm_df_plot2_T2, surv_df_plot_T2)

notation_map <- data.frame(
  Metric = c("pycox.Ant", "pycox.Adj.Ant", 
             "pec", "survC1", "survival.n/G2", "sksurv.ipcw","survival.n", 
             "Hmisc", "SurvMetrics", "lifelines", "pysurvival", 
             "sksurv.censored"),
  Notation = c("C_td", "C_td", 
               "C_tau", "C_tau", "C_tau", 
               "C_tau", "C_tau",
               "C", "C", "C", "C", "C"),
  stringsAsFactors = FALSE
)
df_plot$Metric <- sub("::.*", "", df_plot$Metric)
df_plot <- merge(df_plot, notation_map, by= "Metric", all.x = TRUE)
df_plot$Notation <- factor(df_plot$Notation, levels = c("C", "C_tau", "C_td"),
                           labels = c(
                             "tilde(C)~(Expected~Mortality)",
                             "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
                             "tilde(C)[td]~(Survival~Distribution)"
                           ))
df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit", "CoxPH", "DeepSurv", "CoxTime"))
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.4, color = "black") +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```


```{r}
select_exps <- grep("ExpMort\\.",colnames(subset_expm_B[[1]]), value = TRUE)
results_expm1_B <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm1_B[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("Hmisc::rcorr.cens",
                                                        "pysurvival", 
                                                        "SurvMetrics::Cindex", 
                                                        "lifelines", 
                                                        "sksurv.censored"),
                                  eval.times =
                                    round(max(stacked_predictions_B$test_time)), #C
                                  sampled_data = subset_expm_B)
}

select_exps <- grep("ExpMort\\.",colnames(subset_expm_B[[1]]), value = TRUE)
results_expm2_B <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm2_B[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pec::cindex",
                                                        "survC1::Est.Cval", 
                                                        "survival.n", 
                                                        "survival.n/G2"),
                                  eval.times = round(max(stacked_predictions_B$test_time[stacked_predictions_B$test_status == 1])), # C tau.
                                  sampled_data = subset_expm_B)
}
```



```{r}
model_names <- unique(sub("\\..*", "", grep("^[A-Za-z]+\\.\\d+$",
                                            colnames(subset_surv_B[[1]]), value = TRUE)))
results_surv_B <- list()
for (model in model_names) {
  # Antolinis
  cat("Calculating bootstrap for Distribution for Model = ", model, "\n")
  results_surv_B[[model]] <-  bootstrap.metric(metrics.wrapper,
                                  dataset=list(
                                     predicted = model,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pycox.Ant", "pycox.Adj.Ant"),
                                  sampled_data = subset_surv_B)
}
```


```{r, fig.width=12, fig.height=6}

expm_df_plot1_B <- make_expm_plot_entries(results_expm1_B)
expm_df_plot2_B <- make_expm_plot_entries(results_expm2_B)
surv_df_plot_B <- make_surv_plot_entries(results_surv_B)

df_plot <- bind_rows(expm_df_plot1_B, expm_df_plot2_B, surv_df_plot_B)

notation_map <- data.frame(
  Metric = c("pycox.Ant", "pycox.Adj.Ant", 
             "pec", "survC1", "survival.n/G2", "sksurv.ipcw","survival.n", 
             "Hmisc", "SurvMetrics", "lifelines", "pysurvival", 
             "sksurv.censored"),
  Notation = c("C_td", "C_td", 
               "C_tau", "C_tau", "C_tau", 
               "C_tau", "C_tau",
               "C", "C", "C", "C", "C"),
  stringsAsFactors = FALSE
)
df_plot$Metric <- sub("::.*", "", df_plot$Metric)
df_plot <- merge(df_plot, notation_map, by= "Metric", all.x = TRUE)
df_plot$Notation <- factor(df_plot$Notation, levels = c("C", "C_tau", "C_td"),
                           labels = c(
                             "tilde(C)~(Expected~Mortality)",
                             "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
                             "tilde(C)[td]~(Survival~Distribution)"
                           ))
df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit", "CoxPH", "DeepSurv", "CoxTime"))
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.4, color = "black") +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```

PLotting this small differences...



```{r}
surv_df_plot$Type <- "No rounding"
surv_df_plot_TT$Type <- "Round(T, 1)"
surv_df_plot_T2$Type <- "Round(T, 0)"
surv_df_plot_B$Type <- "Binned (3 months)"

df_all <- bind_rows(surv_df_plot, surv_df_plot_TT, surv_df_plot_T2, surv_df_plot_B)

df_all$Type <- factor(df_all$Type, levels = c("No rounding", "Round(T, 1)", 
                                              "Round(T, 0)", "Binned (3 months)"))

ggplot(df_all, aes(x = Model, y = cindex, shape = Type)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6),
                  size = 0.4, color = "black") +
  facet_wrap(~ Metric, scales = "free_x") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +  # Adjust shapes as needed
  labs(x = NULL, y = "C-index",
       shape = "",
       title = "") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

```

```{r, fig.width=11, fig.height=9}

expm_df_plot1$Type <- " Original Tied Times "
expm_df_plot1_TT$Type <- "Rounded Tied times 1"
expm_df_plot1_T2$Type <- "Rounded Tied Times 2"
expm_df_plot1_B$Type <- "Binned by 3 months"
expm_df_plot2$Type <- " Original Tied Times "
expm_df_plot2_TT$Type <- "Rounded Tied times 1"
expm_df_plot2_T2$Type <- "Rounded Tied Times 2"
expm_df_plot2_B$Type <- "Binned by 3 months"

df_all <- bind_rows(expm_df_plot1, expm_df_plot1_TT, expm_df_plot1_T2, expm_df_plot1_B,
                    expm_df_plot2, expm_df_plot2_TT, expm_df_plot2_T2, expm_df_plot2_B)

ggplot(df_all, aes(x = Model, y = cindex, shape = Type)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6),
                  size = 0.4, color = "black") +
  facet_wrap(~ Metric, scales = "free_x") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +  # Adjust shapes as needed
  labs(x = NULL, y = "C-index",
       shape = "Measurement Type",
       title = "") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```


These are the percentages of tied pairs. The values are very low, that is because we have a total of > a million of pairs. The key to understand this, is that a small number of unique time values with large counts leads to many subjects being tied, but not many pairs.... 

so only large groups create many tied pairs... 


```{r}

calc_perc <- function(stacked_predictions) {
  
  stacked_predictions %>%
    count(test_time) %>%
    ggplot(aes(x = n)) +
    geom_histogram(binwidth = 1, fill = "grey60", color = "black") +
    scale_x_continuous(breaks = seq(1, max(stacked_predictions$test_time), by = 2)) +
    labs(x = "Number of subjects sharing a time", y = "Number of unique times",
         title = "Distribution of Time Frequencies")
  
  time_summary <- stacked_predictions %>%
    count(test_time) %>%
    mutate(tied_pairs = ifelse(n > 1, n * (n - 1) / 2, 0))

  p <- ggplot(time_summary, aes(x = n, y = tied_pairs)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE, color = "blue") +
    labs(x = "Subjects sharing the same time",
         y = "Number of tied pairs",
         title = "Tied Subjects vs Tied Pairs (Quadratic Growth)")
  
  # Total number of tied subjects
  tied_subjects <- stacked_predictions$test_time %in% stacked_predictions$test_time[duplicated(stacked_predictions$test_time)]
  n_tied_subjects <- sum(tied_subjects)
  
  # Total number of tied pairs
  n_tied_pairs <- sum(ifelse(time_summary$n > 1, time_summary$tied_pairs, 0))
  
  # Total number of subjects and pairs
  n_total_subjects <- nrow(stacked_predictions)
  n_total_pairs <- n_total_subjects * (n_total_subjects - 1) / 2
  
  # Percentages
  perc_tied_subjects <- 100 * n_tied_subjects / n_total_subjects
  perc_tied_pairs <- 100 * n_tied_pairs / n_total_pairs
  
  print("% subjects")
  print(perc_tied_subjects)
  print("% tied-time pairs")
  print(perc_tied_pairs)
  print(p)
}

```

```{r}
calc_perc(stacked_predictions = stacked_predictions)
calc_perc(stacked_predictions = stacked_predictions_TT)
calc_perc(stacked_predictions = stacked_predictions_T2)
calc_perc(stacked_predictions = stacked_predictions_B)
```

```{r}
summarize_tied_times <- function(df, time_col = "test_time") {
  df %>%
    count(!!sym(time_col)) %>%                                # Count frequency of each time
    filter(n > 1) %>%                                          # Keep only tied times
    mutate(tied_pairs = (n * (n - 1)) / 2) %>%                 # Manually compute pairs
    group_by(n) %>%                                            # Group by group size
    summarise(
      n_times = n(),                                           # How many times had this group size
      total_pairs = sum(tied_pairs),
      .groups = "drop"
    ) %>%
    arrange(desc(n))                                           # Sort by group size
}

case1_table <- summarize_tied_times(stacked_predictions)
case2_table <- summarize_tied_times(stacked_predictions_TT)
case3_table <- summarize_tied_times(stacked_predictions_T2)
case4_table <- summarize_tied_times(stacked_predictions_B)

case1_table$case <- "Base"
case2_table$case <- "TT"
case3_table$case <- "T2"
case4_table$case <- "Binned"

all_cases <- bind_rows(case1_table, case2_table, case3_table, case4_table)
ggplot(all_cases, aes(x = n, y = total_pairs, fill = case)) +
  geom_col(position = "dodge") +
  scale_x_continuous(breaks = unique(all_cases$n)) +
  labs(x = "Group size (subjects sharing a time)",
       y = "Total tied pairs",
       fill = "Case",
       title = "Distribution of Tied Pairs by Group Size") +
  theme_minimal()

```

Look into the comparable pairs: 


```{r}
surv_ <- extract_batch_metrics(results_surv)
surv_TT <- extract_batch_metrics(results_surv_TT)
surv_T2 <- extract_batch_metrics(results_surv_T2)
surv_B <- extract_batch_metrics(results_surv_B)
```

```{r}
surv_$Type <- "No round"
surv_TT$Type <- "Rounded TT"
surv_T2$Type <- "Rounded T2"
surv_B$Type <- "Binned"

# Combine all into one dataframe
df_all <- bind_rows(surv_, surv_TT, surv_T2, surv_B)

df_long <- df_all %>%
  pivot_longer(
    cols = c("pycox.Ant.CP", "pycox.Adj.Ant.CP", "pycox.Ant.EP", "pycox.Adj.Ant.EP"),
    names_to = "Metric",
    values_to = "Pairs"
  ) %>%
  mutate(
    PairType = case_when(
      grepl("CP", Metric) ~ "Concordant Pairs",
      grepl("EP", Metric) ~ "Comparable Pairs"
  )
)

ggplot(df_long, aes(x = Model, y = Pairs, fill = PairType)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  facet_grid(~Type, scales = "free_y") +
  theme_minimal(base_size = 14) +
  labs(x = "Model", y = "Number of Pairs",
       fill = "Implementation",
       title = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Ties in predictions


```{r}
#stacked_predictions 

# round to 1 
stacked_predictions_TP <- stacked_predictions
stacked_predictions_TP[, 5:ncol(stacked_predictions)] <- round(stacked_predictions[, 5:ncol(stacked_predictions)], 2)

# round to 0
stacked_predictions_TP2 <- stacked_predictions
stacked_predictions_TP2[, 5:ncol(stacked_predictions)] <- round(stacked_predictions[, 5:ncol(stacked_predictions)], 1)

```


```{r}
cat("Number of patients: ", nrow(stacked_predictions), "\n")
cat("Number of unique predictions: ", length(unique(stacked_predictions$ExpMort.CoxPH)), "\n")

table_pred <- table(stacked_predictions$ExpMort.CoxPH)
ties_pred <- table_pred[table_pred  > 1]

num_tied_pairs <- sum(choose(table_pred , 2))
n_tied_pairs <- sum((table_pred * (table_pred  - 1)) / 2)
cat("Number of pairs that have tied preds Cox PH (exp mort): ", num_tied_pairs, "\n")
```


```{r}
cat("Number of patients: ", nrow(stacked_predictions_TP), "\n")
cat("Number of unique predictions: ", length(unique(stacked_predictions_TP$ExpMort.CoxPH)), "\n")

table_pred <- table(stacked_predictions_TP$ExpMort.CoxPH)
ties_pred <- table_pred[table_pred  > 1]

num_tied_pairs <- sum(choose(table_pred , 2))
n_tied_pairs <- sum((table_pred * (table_pred  - 1)) / 2)
cat("Number of pairs that have tied preds Cox PH (exp mort): ", num_tied_pairs, "\n")
```


```{r}
cat("Number of patients: ", nrow(stacked_predictions_TP2), "\n")
cat("Number of unique predictions: ", length(unique(stacked_predictions_TP2$ExpMort.CoxPH)), "\n")

table_pred <- table(stacked_predictions_TP2$ExpMort.CoxPH)
ties_pred <- table_pred[table_pred  > 1]

num_tied_pairs <- sum(choose(table_pred , 2))
n_tied_pairs <- sum((table_pred * (table_pred  - 1)) / 2)
cat("Number of pairs that have tied preds Cox PH (exp mort): ", num_tied_pairs, "\n")
```

```{r, eval=FALSE}
# TEST

set.seed(123)

#subset_stacked <- stacked_predictions[(stacked_predictions$cv_repeat == 1),]

# Number of bootstrapps
n_bootstraps <- 100
# Size of boostrapped sample
size_sample <- 1000

# re sample by patient id
resample_indices <- replicate(n_bootstraps, 
                              sample(stacked_predictions_TP$patients_ids, 
                                     size = size_sample, replace = TRUE), 
                              simplify = FALSE)
```



```{r}
# subset_risk_10_TT <- list()
subset_expm_TP <- list()
subset_surv_TP <- list()

for (r in seq_along(resample_indices)) {
  # subset_risk_10_TT[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TT, 
  #                model_names = "all",
  #                input_type = "RiskAtT",
  #                specific_time = c(60), # 5 years
  #                bootstrap_patient_ids = resample_indices[[r]])
  subset_expm_TP[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TP, 
                 model_names = "all",
                 input_type = "ExpectedMortality",
                 bootstrap_patient_ids = resample_indices[[r]])
  subset_surv_TP[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TP, 
                model_names = "all",
                input_type = "Distribution",
                bootstrap_patient_ids = resample_indices[[r]])
  
  
}
```

```{r}
subset_expm_TP2 <- list()
subset_surv_TP2 <- list()

for (r in seq_along(resample_indices)) {
  # subset_risk_10_TT[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TT, 
  #                model_names = "all",
  #                input_type = "RiskAtT",
  #                specific_time = c(60), # 5 years
  #                bootstrap_patient_ids = resample_indices[[r]])
  subset_expm_TP2[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TP2, 
                 model_names = "all",
                 input_type = "ExpectedMortality",
                 bootstrap_patient_ids = resample_indices[[r]])
  subset_surv_TP2[[r]] <- get_model_preds2(stacked_predictions = stacked_predictions_TP2, 
                model_names = "all",
                input_type = "Distribution",
                bootstrap_patient_ids = resample_indices[[r]])
  
  
}
```


```{r}
select_exps <- grep("ExpMort\\.",colnames(subset_expm_TP[[1]]), value = TRUE)
results_expm1_TP <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm1_TP[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("Hmisc::rcorr.cens",
                                                        "pysurvival",
                                                        "SurvMetrics::Cindex", 
                                                        "lifelines", 
                                                        "sksurv.censored"),
                                  eval.times = round(max(mb$time)), #C
                                  sampled_data = subset_expm_TP)
}

select_exps <- grep("ExpMort\\.",colnames(subset_expm_TP[[1]]), value = TRUE)
results_expm2_TP <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm2_TP[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pec::cindex",
                                                        "survC1::Est.Cval", 
                                                        "survival.n", 
                                                        "survival.n/G2"),
                                  eval.times = round(max(mb$time[mb$status == 1])), # C tau.
                                  sampled_data = subset_expm_TP)
}

model_names <- unique(sub("\\..*", "", grep("^[A-Za-z]+\\.\\d+$",
                                            colnames(subset_surv_TP[[1]]), value = TRUE)))
results_surv_TP <- list()
for (model in model_names) {
  # Antolinis
  cat("Calculating bootstrap for Distribution for Model = ", model, "\n")
  results_surv_TP[[model]] <-  bootstrap.metric(metrics.wrapper,
                                  dataset=list(
                                     predicted = model,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pycox.Ant", "pycox.Adj.Ant"),
                                  sampled_data = subset_surv_TP)
}
```


```{r, fig.width=12, fig.height=6}

expm_df_plot1_TP <- make_expm_plot_entries(results_expm1_TP)
expm_df_plot2_TP <- make_expm_plot_entries(results_expm2_TP)
surv_df_plot_TP <- make_surv_plot_entries(results_surv_TP)

df_plot <- bind_rows(expm_df_plot1_TP, expm_df_plot2_TP, surv_df_plot_TP)

notation_map <- data.frame(
  Metric = c("pycox.Ant", "pycox.Adj.Ant", 
             "pec", "survC1", "survival.n/G2", "sksurv.ipcw","survival.n", 
             "Hmisc", "SurvMetrics", "lifelines", "pysurvival", 
             "sksurv.censored"),
  Notation = c("C_td", "C_td", 
               "C_tau", "C_tau", "C_tau", 
               "C_tau", "C_tau",
               "C", "C", "C", "C", "C"),
  stringsAsFactors = FALSE
)
df_plot$Metric <- sub("::.*", "", df_plot$Metric)
df_plot <- merge(df_plot, notation_map, by= "Metric", all.x = TRUE)
df_plot$Notation <- factor(df_plot$Notation, levels = c("C", "C_tau", "C_td"),
                           labels = c(
                             "tilde(C)~(Expected~Mortality)",
                             "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
                             "tilde(C)[td]~(Survival~Distribution)"
                           ))
df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit", "CoxPH", "DeepSurv", "CoxTime"))
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.4, color = "black") +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```


#### Repeat with more ties


```{r}
select_exps <- grep("ExpMort\\.",colnames(subset_expm_TP2[[1]]), value = TRUE)
results_expm1_TP2 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm1_TP2[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("Hmisc::rcorr.cens",
                                                        "pysurvival", 
                                                        "SurvMetrics::Cindex", 
                                                        "lifelines", 
                                                        "sksurv.censored"),
                                  eval.times = round(max(mb$time)), #C
                                  sampled_data = subset_expm_TP2)
}

select_exps <- grep("ExpMort\\.",colnames(subset_expm_TP2[[1]]), value = TRUE)
results_expm2_TP2 <- list()
for (exps in select_exps) {
   cat("Calculating bootstrap for ExpMort for Model = ", exps, "\n")
   results_expm2_TP2[[exps]] <- bootstrap.metric.parallel(metrics.wrapper,
                                  dataset=list(
                                     predicted = exps,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pec::cindex",
                                                        "survC1::Est.Cval", 
                                                        "survival.n", 
                                                        "survival.n/G2"),
                                  eval.times = round(max(mb$time[mb$status == 1])), # C tau.
                                  sampled_data = subset_expm_TP2)
}
```



```{r}
model_names <- unique(sub("\\..*", "", grep("^[A-Za-z]+\\.\\d+$",
                                            colnames(subset_surv_TP2[[1]]), value = TRUE)))
results_surv_TP2 <- list()
for (model in model_names) {
  # Antolinis
  cat("Calculating bootstrap for Distribution for Model = ", model, "\n")
  results_surv_TP2[[model]] <-  bootstrap.metric(metrics.wrapper,
                                  dataset=list(
                                     predicted = model,
                                     censoring = "test_status",
                                     time = "test_time"),
                                  implementation = list("pycox.Ant", "pycox.Adj.Ant"),
                                  sampled_data = subset_surv_TP2)
}
```


```{r, fig.width=12, fig.height=6}

expm_df_plot1_TP2 <- make_expm_plot_entries(results_expm1_TP2)
expm_df_plot2_TP2 <- make_expm_plot_entries(results_expm2_TP2)
surv_df_plot_TP2 <- make_surv_plot_entries(results_surv_TP2)

df_plot <- bind_rows(expm_df_plot1_TP2, expm_df_plot2_TP2, surv_df_plot_TP2)

notation_map <- data.frame(
  Metric = c("pycox.Ant", "pycox.Adj.Ant", 
             "pec", "survC1", "survival.n/G2", "sksurv.ipcw","survival.n", 
             "Hmisc", "SurvMetrics", "lifelines", "pysurvival", 
             "sksurv.censored"),
  Notation = c("C_td", "C_td", 
               "C_tau", "C_tau", "C_tau", 
               "C_tau", "C_tau",
               "C", "C", "C", "C", "C"),
  stringsAsFactors = FALSE
)
df_plot$Metric <- sub("::.*", "", df_plot$Metric)
df_plot <- merge(df_plot, notation_map, by= "Metric", all.x = TRUE)
df_plot$Notation <- factor(df_plot$Notation, levels = c("C", "C_tau", "C_td"),
                           labels = c(
                             "tilde(C)~(Expected~Mortality)",
                             "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
                             "tilde(C)[td]~(Survival~Distribution)"
                           ))
df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit", "CoxPH", "DeepSurv", "CoxTime"))
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.4, color = "black") +
   facet_wrap(~ Notation, nrow = 1, scales = "free_x", labeller = label_parsed)+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```

All the expected mortality together

```{r, fig.width=12, fig.height=6}

## add type
expm_df_plot1$Type <- "Original"
expm_df_plot1_TP$Type <- "Ties Pred First Rounding"
expm_df_plot1_TP2$Type <- "Ties Pred Second Rounding"

df_plot <- bind_rows(expm_df_plot1, expm_df_plot1_TP, expm_df_plot1_TP2)

# Merge to get Notation.map (new column for matching from map)
df_plot <- merge(df_plot, notation_map, by = "Metric", all.x = TRUE, suffixes = c("", ".map"))

# Update Notation column only if it's missing (i.e., NA)
df_plot$Notation <- ifelse(is.na(df_plot$Notation), df_plot$Notation.map, df_plot$Notation)

# Remove the temporary map column
df_plot$Notation.map <- NULL

# Re-factor Notation for plotting
df_plot$Notation <- factor(df_plot$Notation, levels = c("C", "C_tau", "C_td"),
                           labels = c(
                             "tilde(C)~(Expected~Mortality)",
                             "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
                             "tilde(C)[td]~(Survival~Distribution)"
                           ))

df_plot$Metric <- sub("::.*", "", df_plot$Metric)

df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit", "CoxPH", "DeepSurv", "CoxTime"))
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Type, nrow = 1, scales = "free_x")+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.4, color = "black") +
   facet_wrap(~ Type, nrow = 1, scales = "free_x")+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```

```{r, fig.width=12, fig.height=6}

## add type
expm_df_plot2$Type <- "Original"
expm_df_plot2_TP$Type <- "Ties Pred First Rounding"
expm_df_plot2_TP2$Type <- "Ties Pred Second Rounding"

df_plot <- bind_rows(expm_df_plot2, expm_df_plot2_TP, expm_df_plot2_TP2) 

# Merge to get Notation.map (new column for matching from map)
df_plot <- merge(df_plot, notation_map, by = "Metric", all.x = TRUE, suffixes = c("", ".map"))

# Update Notation column only if it's missing (i.e., NA)
df_plot$Notation <- ifelse(is.na(df_plot$Notation), df_plot$Notation.map, df_plot$Notation)

# Remove the temporary map column
df_plot$Notation.map <- NULL

# Re-factor Notation for plotting
df_plot$Notation <- factor(df_plot$Notation, levels = c("C", "C_tau", "C_td"),
                           labels = c(
                             "tilde(C)~(Expected~Mortality)",
                             "tilde(C)[tau]~(Expected~Mortality~tau==max*(T:~Delta==1))",
                             "tilde(C)[td]~(Survival~Distribution)"
                           ))

df_plot$Metric <- sub("::.*", "", df_plot$Metric)

df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit", "CoxPH", "DeepSurv", "CoxTime"))
df_plot$Type <- factor(df_plot$Type)
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Type, nrow = 1, scales = "free_x")+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.4, color = "black") +
   facet_wrap(~ Type, nrow = 1, scales = "free_x")+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```

```{r, fig.width=12, fig.height=7}

## add type
surv_df_plot$Type <- "Absence of rounding"
surv_df_plot_TP$Type <- "Round(M(x), 2)"
surv_df_plot_TP2$Type <- "Round(M(x), 1)"

df_plot <- bind_rows(surv_df_plot, surv_df_plot_TP, surv_df_plot_TP2)
# Merge to get Notation.map (new column for matching from map)

df_plot$Model <- factor(df_plot$Model, levels = c("RSF", "DeepHit",
                                                  "CoxPH", "DeepSurv",
                                                  "CoxTime"))

df_plot$Type <- factor(df_plot$Type, levels = c("Absence of rounding",
                                                "Round(M(x), 2)", 
                                                "Round(M(x), 1)"))
### Version with color
 ggplot(df_plot, aes(x = Metric, y = cindex, color = Model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.6), size = 0.4) +
   facet_wrap(~ Type, nrow = 1, scales = "free_x")+
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  labs(title = "",
       y = "C-index", x = NULL) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
 
### Version for paper
p <- ggplot(df_plot, aes(x = Metric, y = cindex)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, shape=Model),
                  position = position_dodge(width = 0.6), size = 0.6, color = "black") +
   facet_wrap(~ Type, nrow = 1, scales = "free_x")+
   scale_shape_manual(values = c("RSF" = 15, "DeepHit" = 23, 
                                 "CoxPH" = 19, "DeepSurv" = 22, "CoxTime" = 24)) +
   geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
   labs(title = "",
       y = "C-index", x = NULL) +
   theme_minimal(base_size = 19) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

print(p)

#ggsave("./Results/PaperTables/ExpectedMortalityBlack.png", plot = p, width = 12, height = 6, dpi = 300)
```


